---
title: "RAMP Country-Device Analyses"
author: "Minh Pham, Nikolaus Parulian, Jon Wheeler"
date: "3/2/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

These are the calculations used to generate the tables and figures in the manuscript.

File paths for data file imports are relative to the structure of the GitHub repository. Update file paths as needed. If this notebook was downloaded with the GitHub repository, it should work as-is.

Packages imported in the next code block can be installed from CRAN as needed.

```{r}
setwd('.')
library(rio)
library(dplyr)
library(psych)
```
### Page Click Analysis

The article includes two different analyses of RAMP data. The first analysis uses Google Search Console (GSC) API

```
dat1 <- import("../ramp_data/2019-01_RAMP_subset_country-device-info.csv")
dat2 <- import("../ramp_data/2019-02_RAMP_subset_country-device-info.csv")
dat3 <- import("../ramp_data/2019-03_RAMP_subset_country-device-info.csv")
dat4 <- import("../ramp_data/2019-04_RAMP_subset_country-device-info.csv")
dat5 <- import("../ramp_data/2019-05_RAMP_subset_country-device-info.csv")
str(dat1)
```
```{r}
str(dat2)
```
```{r}
str(dat3)
```
```{r}
str(dat4)
```
```{r}
str(dat5)
```
```{r}
dat1 <- dat1[, c(-2, -5, -6)]
dat2 <- dat2[, c(-2, -5, -6)]
dat3 <- dat3[, c(-2, -5, -6)]
dat4 <- dat4[, c(-2, -5, -6)]
dat5 <- dat5[, c(-2, -5, -6)]
country_device <- rbind(dat1, dat2)
country_device <- rbind(country_device, dat3)
country_device <- rbind(country_device, dat4)
country_device <- rbind(country_device, dat5)
length(unique(country_device$index))
```
```{r}
country_device <- country_device[, c(-5)]
```
```{r}
length(unique(country_device$country))
```

The next two code blocks are for table 11 and figure 10.

```{r}
country_device %>% 
  select(clicks, device) %>% 
  group_by(device) %>% 
  summarise(clicks = sum(clicks)) %>% 
  arrange(desc(clicks)) %>% 
  mutate(percent = clicks/sum(clicks)*100)
```
Here's some code that reproduces the data we have in table 11 now:

```{r}
country_device %>% 
  select(clicks, device) %>% 
  group_by(device) %>% 
  summarise(device_row_count = sum(!is.na(clicks))) %>% 
  arrange(desc(device_row_count)) %>% 
  mutate(percent = device_row_count/4333430*100)
```

```{r}
dat6 <- import('../ir_data/North_South.xlsx')
head(dat6)
```
```{r}
length(dat6$Countryabbre)
```
```{r}
dat6$Countryabbre <- tolower(dat6$Countryabbre)
head(dat6)
```
```{r}
country_device_n <- merge(country_device, dat6, by.x = "country", by.y = "Countryabbre")
length(unique(country_device_n$country))
```
```{r}
dat7 <- unique(country_device$country)
dat8 <- unique(dat6$Countryabbre)
dat7[!(dat7 %in% dat8)]
```
```{r}
dat8[!(dat8 %in% dat7)]
```
```{r}
write.csv(country_device_n, file = '../ir_data/country_device_n.csv')
```
```{r}
describeBy(country_device_n$clicks, country_device_n$device)
```

```{r}
describeBy(country_device_n$position, country_device_n$device)
```
```{r}
country_device_n %>% select(clicks, country) %>% 
  group_by(country) %>% 
  summarise(clicks = sum(clicks)) %>% 
  arrange(desc(clicks)) %>% 
  head(5)
```


```{r}
country_device_n$Location <- as.factor(country_device_n$Location)
```
```{r}
country_device_n %>% 
  select(clicks, Location) %>% 
  group_by(Location) %>% 
  summarise(clicks = sum(clicks)) %>% 
  arrange(desc(clicks)) %>% 
  mutate(percent = clicks/sum(clicks)*100)
```
```{r}
country_device_n%>%
  select(clicks, device, Location)%>%
  group_by(Location, device)%>%
  summarise(clicks = sum(clicks))%>%
  arrange(desc(clicks))%>%
  mutate(percent = clicks/sum(clicks)*100)
```
```{r}
mean(country_device_n$clicks)
```
```{r}
var(country_device_n$clicks)
```
```{r}
range(country_device_n$clicks)
```
```{r}
mod <- glm(clicks~position*device, data = country_device_n, family = c("quasipoisson"))
summary(mod)
```
```{r}
round(cbind(exp(coef(mod)),
            exp(confint(mod))),3)
```