---
title: "RAMP Country-Device Analyses"
author: "Minh Pham, Nikolaus Parulian, Jon Wheeler"
date: "3/2/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

GitHub repository URL: <https://github.com/imls-measuring-up/ramp-analyses-scripts>

## R Markdown

These are the calculations used to generate the tables in [the manuscript].

The figures use the same summary data as the analyses presented here. This summary data is included in the "../results/" directory. The filename is "RAMP_summary_stats_20200109.csv." The figures were generated by Tableau and are available through a publicly accessible dashboard at <...>.

File paths for data file imports are relative to the structure of the GitHub repository. Update file paths as needed. If this notebook was downloaded with the GitHub repository, it should work as-is.

Packages imported in the next code block can be installed from CRAN as needed.

```{r load_packages, echo=TRUE}
setwd('.')
library(rio)
library(dplyr)
library(psych)
```
### About the Data

[The article] includes two different analyses of RAMP data. Both use search engine performance data about RAMP IR harvested from the Google Search Console (GSC) API, though the data used for each analysis are different. For detailed information about how RAMP data are harvested, processed, and exported to CSV format, please see documentation included with the published subset of RAMP data:

> Wheeler, Jonathan et al. (2020), RAMP data subset, January 1 through May 31, 2019, v6, University of New Mexico, Dataset, <https://doi.org/10.5061/dryad.fbg79cnr0>

The RAMP data published as cited above are the same as the data that were used for the analyses in the article. For the reported analyses, the data were further augmented with some additional data about indvidual RAMP IR and global north-south designations of countries represented in the search engine performance data harvested from GSC. All data, data aggregation, and analysis scripts necessary to reproduce the reported analyses and findings are included, with documentation, in the the GitHub repository that includes this R Markdown file.

Section headings provided here are the same as used in the article. Where code blocks output results used in the article, those code blocks are labeled with the corresponding figure or table number.

### Materials and Methods

#### Table 1

Table 1 is a subset of columns from the manually tabulated information about IR size and platform. These data are included in the "RAMP_IR_base_info.csv" file, which is in the "ir_data" directory in the GitHub repository. 

```{r import_subset_ir_info_table_1, echo=TRUE}
ir_platform_size <- import("../ir_data/RAMP_IR_base_info.csv")

ir_platform_size %>%
  select("Country", "Type", "Repository Name",
         "Normalized_Platform", "Items in repository on 2019-05-27") %>% 
  group_by(Country, Type) %>% 
  arrange(desc(Country), desc(Type))

```


### Page Click Analysis

This analysis uses information about RAMP IR that was manually gathered in May and June 2019. For each IR included in the study, this information is cross referenced with aggregated page-click data to provide summary statistics about IR search engine performance. There are limitations to the conclusions that may be drawn from this summary analysis - please see the article for more information. **In particular, the authors of this study strongly discourage qualitative assessments of any and all IR included in this study. The results provided by this analysis, as reported, should not be taken as an assessment of the quality of IR providers, platforms, and/or content.**

The summary statistics used for the reported analysis were generated on January 9, 2020. The output is included in the GitHub repository and is referenced in the corresponding import path below. Note that the script and data needed to reproduce those same summary stats are included in this GitHub repository. Interested parties are invited to redo the aggregation and run this notebook against the new output. If the same data are used, any discrepancies with published results should be brought to the authors' attention.

#### Table 2

```{r import_RAMP_summary_stats, echo=TRUE}
## Read the csv file with descriptive statistics for RAMP IR.
## Note the filename incluces a date, which may need to be updated.

## The file referenced in the next line was used for the reported 
## analysis and is included in the GitHub repo. It is not necessary
## to change the path unless you want to run against more recently
## aggregated data.
IR <- import("../results/RAMP_summary_stats_20200109.csv")
## str(IR)
```
```{r table_2}
## Caption: Overall statistics of items in repositories registered with RAMP
## 35 IR were included in the study, so the output of the length() command should be 35.
## The data frame length and the summary data are included in Table 2.
length(IR$ir)
summary(IR$countItems)
```

#### Table 3

```{r table_3, echo=TRUE}
## Caption: Number of items by platform.

IR%>% 
  group_by(normIrPlat)%>% 
  summarise(sum = sum(countItems))%>%
  arrange(desc(sum))
IR$Platform <- factor(IR$normIrPlat, levels = c("DSpace", "Digital Commons", "EPrints", "Fedora"))
describeBy(IR$countItems, IR$Platform)
```

#### Table 4

```{r table_4, echo=TRUE}
## Caption: Number of repository items per country
## Item counts summarized by country where the IR are located.
IR%>% 
  group_by(irCountry)%>% 
  summarise(sum = sum(countItems))%>%
  arrange(desc(sum))
IR$Country <- factor(IR$irCountry, levels = c("USA", "Australia", "UK", "Canada", "Sweden", "New Zealand", "South Africa"))
describeBy(IR$countItems, IR$irCountry)

```

The table below describes IR item counts grouped by host organization type. The types are "University" and "Consortium." This table is not included in the manuscript, but is left in this notebook for interest's sake.

```{r item_counts_by_org_type, echo=TRUE}
## Item counts by Type of Host Organization - University of Consortium
IR%>% 
  group_by(irType)%>% 
  summarise(sum = sum(countItems))%>%
  arrange(desc(sum))
IR$Type <- factor(IR$irType, levels = c("University", "Consortium"))
describeBy(IR$countItems, IR$irType)
```


#### Table 5

This first code block outputs the value of "N" (count of repositories) for the use ratio categories in Table 5.

```{r table_5_n, echo=TRUE}
## "Caption Number of items by use ratio category"
## Count of IR in each use ratio category.
## Create use ratio category. Item count by category.
IR$category <- ifelse(IR$useRatio<=.5 & IR$countItems > 51000, "High Item Low Downloads",
                      ifelse(IR$useRatio>=.51 & IR$countItems<= 17979,"Low items High Downloads", "Low items Low Downloads"))
table(IR$category)
```

This next code block outputs the corresponding descriptive statistics for each use ratio category in Table 5.

```{r table_5_summary_stats, echo=TRUE}
# Descriptive stats by use ratio category
IR$category <- factor(IR$category, levels = c("High Item Low Downloads", "Low items Low Downloads", "Low items High Downloads"))
describeBy(IR$countItems, IR$category)
```

This code block outputs the total number of items corresponding to each use ratio category in Table 5.

```{r table_5_totals, echo=TRUE}
## Total number of items in each use ratio category.
IR%>% 
  group_by(category)%>% 
  summarise(sum = sum(countItems))%>%
  arrange(desc(sum))
```

#### Table 6

```{r table_6, echo=TRUE}
# Caption "Use ratio category"
describeBy(IR$useRatio, IR$category)
```

#### Table 7

```{r table_7, echo=TRUE}
# Use ratio per platform
IR$normIrPlat <- factor(IR$normIrPlat, levels = c("EPrints", "Fedora", "DSpace", "Digital Commons"))
describeBy(IR$useRatio, IR$normIrPlat)
```

#### About Table 8

Table 8 is a representation of the information in the file *RAMP_summary_stats_20200109.csv*, ordered by use ratio. The table in the article is formatted somewhat differently, but the following code block provides the necessary logic.

```{r table_10, echo=TRUE}
IR %>% 
  select(repoName, countItems, countItemUrls, useRatio) %>% 
  arrange(desc(useRatio))
```

### Country-Device Analysis

Info here.

```{r import_country_device_RAMP_data, echo=TRUE}
dat1 <- import("../ramp_data/2019-01_RAMP_subset_country-device-info.csv")
dat2 <- import("../ramp_data/2019-02_RAMP_subset_country-device-info.csv")
dat3 <- import("../ramp_data/2019-03_RAMP_subset_country-device-info.csv")
dat4 <- import("../ramp_data/2019-04_RAMP_subset_country-device-info.csv")
dat5 <- import("../ramp_data/2019-05_RAMP_subset_country-device-info.csv")
str(dat1)
```
```{r}
str(dat2)
```
```{r}
str(dat3)
```
```{r}
str(dat4)
```
```{r}
str(dat5)
```
```{r}
dat1 <- dat1[, c(-2, -5, -6)]
dat2 <- dat2[, c(-2, -5, -6)]
dat3 <- dat3[, c(-2, -5, -6)]
dat4 <- dat4[, c(-2, -5, -6)]
dat5 <- dat5[, c(-2, -5, -6)]
country_device <- rbind(dat1, dat2)
country_device <- rbind(country_device, dat3)
country_device <- rbind(country_device, dat4)
country_device <- rbind(country_device, dat5)
rm(dat1, dat2, dat3, dat4, dat5)
length(unique(country_device$index))
```

The next two code blocks are for table 11 and figure 10.

```{r table_11, echo=TRUE}
country_device %>% 
  select(clicks, device) %>% 
  group_by(device) %>% 
  summarise(count = sum(!is.na(device)), clicks = sum(clicks)) %>% 
  arrange(desc(clicks)) %>% 
  mutate(percent = clicks/sum(clicks)*100)
```

```{r figure_10, echo=TRUE}
country_device %>% 
  select(clicks, device, index) %>% 
  group_by(index, device) %>% 
  summarise(total_clicks = sum(clicks)) %>% 
  arrange(index, desc(total_clicks)) %>% 
  mutate(pct_clicks_device = total_clicks/sum(total_clicks)*100)
```

Note that this table/figure are not in the current draft. Because of how the country-device data are aggregated by GSC, we cannot assert that a count of occurrences of devices within the data represents a count of _visits_ by device. This is because the data are aggregated by combination of country and device. For example, we may see on any given day that the UNM Digital Repository received a certain number of clicks from desktops users in the United States. However, since these clicks are aggregated by combination of country and device, we don't know how many individual desktop users in the US visited UNM's IR on that day. 

A figure representing the table produced by the next chunk has been removed from the manscript, but the table may be of interest as long as the difference between "occurrence" and "visit" is understood.

```{r figure_10_removed, echo=TRUE}
country_device %>% 
  select(device, index) %>% 
  group_by(index, device) %>% 
  summarise(count_device_occurrence = sum(!is.na(device))) %>% 
  arrange(index, desc(count_device_occurrence)) %>% 
  mutate(pct_access_device = count_device_occurrence/sum(count_device_occurrence)*100)
```

Tables 12 and 13, Figure 11

```{r table_12_clicks, echo=TRUE}
describeBy(country_device$clicks, country_device$device)
```

```{r table_12_position, echo=TRUE}
describeBy(country_device$position, country_device$device)
```

```{r}
mod <- glm(clicks~position*device, data = country_device, family = c("quasipoisson"))
summary(mod)
```
```{r}
round(cbind(exp(coef(mod)),
            exp(confint(mod))),3)
```

## Access by country, and device use across countries from the global north and south

From here on is the code for the global north/south device usage analysis. This requires an additional import of manually tabulated data that merges global north/south designations per country with ISO three letter country codes. 

Data sources to cite:

https://meta.wikimedia.org/wiki/List_of_countries_by_regional_classification

https://en.wikipedia.org/wiki/ISO_3166-1_alpha-3

https://www.iso.org/iso-3166-country-codes.html

Data were manually tabulated/merged by Minh Pham.

```{r, echo=TRUE}
country_device <- country_device[, c(-5)]
```
```{r}
length(unique(country_device$country))
```

```{r}
dat6 <- import('../country_codes/north_south.csv')
head(dat6)
```
```{r}
length(dat6$Countryabbre)
```

Convert country codes to lowercase. They are lowercase in the RAMP data, so this allows us to merge.

```{r}
dat6$Countryabbre <- tolower(dat6$Countryabbre)
head(dat6)
```

This is the merge. Note the count of unique "country" values in the merged dataset is 2 less than it was before the merge. 

```{r}
country_device_n <- merge(country_device, dat6, by.x = "country", by.y = "Countryabbre")
length(unique(country_device_n$country))
```

Look at the difference. First look at country codes from RAMP that are not in the *north_south.csv* file.

The unmatched codes are "zzz" and "xkk," which are (most likely) GSC codes for "unknown region." These data should be dropped from global north/south analyses, since we cannot make a determination of global north/south status for the corresponding search result data.

Dropping these data means that the click sums and row counts by device will not have the same totals in the following analysis as are presented in table 11. Row counts and click sums for these unknown region codes are presented here to document the difference.

```{r}
dat7 <- unique(country_device$country)
dat8 <- unique(dat6$Countryabbre)
dat7[!(dat7 %in% dat8)]
```
```{r}

unknown_region <- country_device %>% filter(country == 'zzz' | country == 'xkk')
```

```{r, echo=TRUE}
unknown_region %>% 
  select(clicks, device) %>% 
  group_by(device) %>% 
  summarise(count = sum(!is.na(device)), clicks = sum(clicks)) %>% 
  arrange(desc(clicks)) %>% 
  mutate(percent = clicks/sum(clicks)*100)
```


```{r, echo=TRUE}
country_device_n %>% 
  select(clicks, device) %>% 
  group_by(device) %>% 
  summarise(count = sum(!is.na(device)), clicks = sum(clicks)) %>% 
  arrange(desc(clicks)) %>% 
  mutate(percent = clicks/sum(clicks)*100)
```

Adding the "count" and "clicks" values per device in the two tables above provides the same results as table 11, which indicates that there are no known regions that are unaccounted for in the *country_device_n* dataframe.

Next we identify countries that are in the country code data but not in RAMP. This just means there were no searches conducted from these countries between Jan 1 and May 31, 2019, which returned RAMP IR content in the SERP. So there is nothing to fix here.

```{r}
dat8[!(dat8 %in% dat7)]
```

```{r}
describeBy(country_device_n$clicks, country_device_n$device)
```

```{r}
describeBy(country_device_n$position, country_device_n$device)
```
```{r table_14}
country_device_n %>% select(clicks, country) %>% 
  group_by(country) %>% 
  summarise(clicks = sum(clicks)) %>% 
  arrange(desc(clicks)) %>% 
  head(5)
```


```{r}
country_device_n$Location <- as.factor(country_device_n$Location)
```
```{r table_15}
country_device_n %>% 
  select(clicks, Location) %>% 
  group_by(Location) %>% 
  summarise(clicks = sum(clicks)) %>% 
  arrange(desc(clicks)) %>% 
  mutate(percent = clicks/sum(clicks)*100)
```
```{r}
country_device_n%>%
  select(clicks, device, Location)%>%
  group_by(Location, device)%>%
  summarise(clicks = sum(clicks))%>%
  arrange(desc(clicks))%>%
  mutate(percent = clicks/sum(clicks)*100)
```
```{r}
mean(country_device_n$clicks)
```
```{r}
var(country_device_n$clicks)
```
```{r}
range(country_device_n$clicks)
```